Variants on the Flight and Maintenance Planning problem

### 2018-07-13 latex/notes/complexite/FMP_variants.tex
last changes to complexity analysis

    In order to generalize the formulation, from now on aircraft will be named "resources" and missions will be named "tasks". A series of $j \in \mathcal{J}$ tasks are planned along a horizon divided into $t \in \mathcal{T}$ periods. Since all tasks are already scheduled, we know time periods $T_j \subset \mathcal{T}$ in which they will be realized. Similarly, all the tasks to be realized in period $t$ are known and defined by set $J_t \subset \mathcal{J}$. Each task requires a certain number $r_{j}$ of resources $i \in \mathcal{I}$ which it employs for a time duration defined by $h_j$ in each period. Set $a_{j} \subset \mathcal{I}$ lists the resources that can be assigned to each task and set $O_i \subset \mathcal{J}$ consists of tasks for which resource $i$ can be used. The number of resources in use for each period $d_t$ is also known.

@@

    For each period $k$ from the original problem, we create a period $t \in \mathcal{T}$. We can use the elements in these two sets indistinctly.
    For each period $k$ with demand $b_{k}$, we create a task $j$ with $T_j = \{t\}$ that requires $r_{j} = b_{t}| t \in T_j$ resources. This implies that $|J_t| = 1$ and $d_t = b_{t}$
    For each worker $i$ we have a corresponding resource $i$.
    Since all resources can cover the demand, we assume $O_i = \{j \in \mathcal{J}\}$ and $a_{j} = \{i \in \mathcal{I}\}$.
    All resources start the planning horizon have just exited a maintenance in the previous period. This implies $n_t = 0$.
    The duration of the maintenance will be equal to $M = \overline{D}^{off} = \underline{D}^{off}$. This is assumption is valid because the original proof was done with fixed values instead of ranges.
    The maximum remaining elapsed time after a maintenance is equal to $E = \overline{D}^{off}$. The minimum elapsed time before a maintenance is equal to $e=\underline{D}^{off}$.
    The usage time for each resource by any task will be $h_j = 0$ and the remaining usage time after a maintenance is $H = 0$. The minimum number of maintenances for each resource will be given by $\lceil(|\mathcal{T}| - \overline{D}^{work}) / M \rceil$.

    The model notation follows closely the \cite{Smet2015} 2.6.1 example.

    \begin{itemize}
         \item objective: minimize the number of aircraft to maximize availability.
    \end{itemize}

@@

    \begin{tabular}{ll}
        $t \in \mathcal{T}$ & Planning horizon periods. \\
        $i \in \mathcal{I}$ & Resources. \\
        $j \in \mathcal{J}$ & tasks. 
    \end{tabular}

@@

    \begin{tabular}{ll}
        $m$ & Minimum number of maintenances per resource in planning horizon. \\
        $M$ & Duration of maintenance. \\
        $E$ & Maximum elapsed time before assigning a maintenance. \\
        $e$ & Minimum elapsed time before assigning a maintenance. \\
        $r_{j}$ & Need of resources for task $j$. \\
    \end{tabular}

    (here we could, in theory just copy the same model than we had in MOSIM).

@@

    % \subsection{Variables}

    % \begin{itemize}
    %     \item $x_{it}$: resource $i$ is working on period $t$.
    %     \item $y_{i}$: resource $i$ is working on at least one period.
    % \end{itemize}

    % \subsection{Model}

    % \begin{align}
    %     & \text{Min}\; \sum_{i \in \mathcal{I}} y_{i} \label{min_num_resources}
    % \end{align}

@@

>>>>

    In order to generalize the formulation, from now on aircraft will be named "resources" and missions will be named "tasks". A series of $j \in \mathcal{J}$ tasks are planned along a horizon divided into $t \in \mathcal{T}$ periods. Since all tasks are already scheduled, we know time periods $T_j \subset \mathcal{T}$ in which they will be realized. Similarly, all the tasks to be realized in period $t$ are known and defined by set $J_t \subset \mathcal{J}$. Each task requires a certain number $R_{j}$ of resources $i \in \mathcal{I}$ which it employs for a time duration defined by $H_j$ in each period. Set $I_{j} \subset \mathcal{I}$ lists the resources that can be assigned to each task and set $O_i \subset \mathcal{J}$ consists of tasks for which resource $i$ can be used. The number of resources in use for each period $D_t$ is also known.

@@

    For each period $k$ from the original problem, we create a period $t \in \mathcal{T}$. We can use the elements in these two sets indistinctly.
    For each period $k$ with demand $b_{k}$, we create a task $j$ with $T_j = \{t\}$ that requires $R_{j} = b_{t}| t \in T_j$ resources. This implies that $J_t = \{j\}| t \in T_j$, $|J_t|=1$ and $D_t = b_{t}$.

    For each worker $i$ we have a corresponding resource $i$.
    Since all resources can cover the demand, we assume $O_i = \{j \in \mathcal{J}\}$.
    All resources start the planning horizon have just exited a maintenance in the previous period. This implies $N_t = 0$.
    The duration of the maintenance will be equal to $M = \overline{D}^{off} = \underline{D}^{off}$. This assumption is valid because the original proof was done with fixed values instead of ranges.

    The maximum remaining elapsed time after a maintenance is equal to $E = \overline{D}^{off}$. The minimum elapsed time before a maintenance is equal to $e=\underline{D}^{off}$.
    The usage time for each resource by any task will be $H_j = 0$ and the remaining usage time after a maintenance is $H = 0$. The minimum number of maintenances for each resource will be given by $M_{min}=\lceil(|\mathcal{T}| - \overline{D}^{work}) / M \rceil$. (Alternatively, the limit could be measured as a minimum number of maintenance months $M_{min}=|\mathcal{T}| - \overline{D}^{work}$).

    % \subsection{Hypothesis}

    % \begin{itemize}
    %      \item objective: minimize the number of aircraft to maximize availability.
    % \end{itemize}

@@

    \begin{tabular}{ll}
        $\mathcal{T}$ & Planning horizon periods. \\
        $\mathcal{I}$ & Resources. \\
        $\mathcal{J}$ & tasks.  \\
        $\mathcal{T}_j$ &  time periods $t \in \mathcal{T}$ in which task $j$ is active. \\
        $\mathcal{J}_t $    &  tasks $j \in \mathcal{J}$ to be realized in period $t$. \\
        $\mathcal{I}_j$     &  resources $i \in \mathcal{I}$ that can be assigned to task $j$. \\
        $\mathcal{O}_i$     &  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. \\
        $\mathcal{T}^{s}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}$\}. \\
    \end{tabular}

@@

    \begin{tabular}{ll}
        $M_{min}$ & Minimum number of maintenances per resource in planning horizon. \\
        $M$ & Duration of maintenance. \\
        $E$ & Maximum elapsed time before assigning a maintenance. \\
        $E_{min}$ & Minimum elapsed time before assigning a maintenance. \\
        $R_{j}$ & Need of resources for task $j$. \\
    \end{tabular}

    % (here we could, in theory just copy the same model than we had in MOSIM).

@@

        \begin{tabular}{ll}
            $m_{it}$    &  =1 if resource $i \in I$ starts a maintenance operation in period $t \in \mathcal{T}$, 0 otherwise. \\
            $ret_{it}$  &  remaining elapsed time (integer) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. \\  
        \end{tabular}

        The following modifications need to be done to the original problem described in MOSIM

        \begin{align}
            & ret_{it} \in [E_{min},E] 
                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:me}\\
            & \sum_{t \in \mathcal{T}} m_{it} \geq M_{min} 
                & i \in \mathcal{I} \label{eq:min_maintenace}
        \end{align}

        Constraints (\ref{eq:me}), which already existed, now have a specific minimum value instead of just 0. Constraints (\ref{eq:min_maintenace}) are new. They put a minimum number of maintenances for each resource. This had been considered as a possible cut to the original problem and does not imply a big change to the solutions.

        The alternative to constraints (\ref{eq:min_maintenace}), in case we want to control the months under maintenance and not the number of maintenance operations, are constraints (\ref{eq:min_maintenace2}).

        \begin{align}
            & \sum_{t \in \mathcal{T}} \sum_{t' \in \mathcal{T}^{s}_t} m_{it'} \geq M_{min}
                & i \in \mathcal{I} \label{eq:min_maintenace2}\\
        \end{align}

        % \begin{align}
        %     % maximum capacity1:
        %     & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \leq m_{max}
        %       & t \in \mathcal{T} \label{eq:capacity1}\\
        %    %avail
        %    & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t + D_t\leq u_{max} 
        %     &t \in \mathcal{T} \label{eq:avalaibility1}\\
        %    % maximum capacity2:                
        %     & \sum_{i \in \mathcal{I}_j} a_{jti} = R_j
        %             & j \in \mathcal{J}, t \in \mathcal{T}_j  \label{eq:taskres}\\
        %     & \sum_{t' \in \mathcal{T}^{s}_t} m_{it'} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} \leq 1 
        %             & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}
        % \end{align}

        % \begin{align}
        %     % remaining used time
        %      & rut_{it} \leq rut_{it-1} + H m_{it} - \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j \notag \\
        %             && t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rut_upper}\\
        %     & rut_{i0} = Rut^{Init}_i
        %            & i \in \mathcal{I} \label{eq:rut_initial}\\
        %     & rut_{it} \geq H m_{it}
        %             & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:rut_lower}\\ 
        %     & rut_{it} \in [0,H]
        %             & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:mu} \\
        %     & ret_{it} \leq ret_{it-1} - 1 + E m_{it}
        %             & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:ret_upper}\\
        %     & ret_{i0} = Ret^{Init}_i
        %             & i \in \mathcal{I} \label{eq:ret_initial}\\
        %     & ret_{it} \geq E m_{it}
        %             & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_lower}\\                 
        %     & ret_{it} \in [0,E]
        %             & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:me}\\
        %     & \sum_{i \in \mathcal{I}} ret_{it} \geq Ret^{Init}_{sum}
        %           & t = |\mathcal{T}| \label{eq:min_ret}\\
        %     & \sum_{i \in \mathcal{I}} rut_{it} \geq Rut^{Init}_{sum}
        %           & t = |\mathcal{T}| \label{eq:min_rut}
        % \end{align}
    % \begin{align}
    %     & \text{Min}\; \sum_{i \in \mathcal{I}} y_{i} \label{min_num_resources}
    % \end{align}

@@

\section{Fixed interval scheduling}

    From \cite{Smet2015}, \cite{Kolen2007} and \cite{Kovalyov2007}.

    \cite{Smet2015} calls it "shift minimization personnel task scheduling problem. When it's not related to workers and in a more general way, it's usually called "Fixed Interval Scheduling Problem".

    Here is the description of \cite{Smet2015}:

    Let $J = {1, ..., n}$ be the set of tasks to be assigned and $E = {1, ..., m}$ the set of employees. Each task $j \in J$ has a duration $u_j$ , a start time $s_j$ and a finish time $f_j$ = $s_j + u_j$ . Each employee e has a set of tasks $J_e \in J$ that he/she can perform. Similarly, for each task $j$, a set $E_j \in E$ exists, which contains all employees that can perform task $j$. Both $T_e$ and $E_j$ are defined based on qualifications, time windows of tasks and availabilities of employees.

    For each employee $e \in E$, we create an analogous resource $i \in I$ which we will use indistinctly.

    For each task $j \in J$, we create a $j \in J$ task in our problem with duration and minimal assignment time equal to $A^{min}_j = u_j$. Start times and finish times define the moment when the task is active: $\mathcal{T}_j = t \in \{s_j ... f_j\}$. The need of resources for each task will be constant $R_{j} = 1$. The number of resources working at any given time can $D_t = |\mathcal{J}_t|$ can be calculated. The relationship between tasks and employees is also equivalent: $\mathcal{O}_i = J_i$ and $\mathcal{I}_j = E_j$.

    The usage time for each resource by any task will be $H_j = 0$ and the remaining usage time after a maintenance is $H = 0$.
    For maintenances, we can just take them out of the equation completely. For this we just need to put: $E_{min} = E = M = N_t = 0$ and $Ret^{Init}_{i} = |\mathcal{T}| + 1$.

    \begin{tabular}{ll}
        $\mathcal{T}$       & Planning horizon periods. \\
        $\mathcal{I}$       & Resources. \\
        $\mathcal{J}$       & tasks.  \\
        $\mathcal{T}_j$     &  time periods $t \in \mathcal{T}$ in which task $j$ is active. \\
        $\mathcal{J}_t$     &  tasks $j \in \mathcal{J}$ to be realized in period $t$. \\
        $\mathcal{I}_j$     &  resources $i \in \mathcal{I}$ that can be assigned to task $j$. \\
        $\mathcal{O}_i$     &  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. \\
        $\mathcal{T}^{s}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}$\}. \\
    \end{tabular}

    \begin{tabular}{ll}
        $M_{min}$        & Minimum number of maintenances per resource in planning horizon. \\
        $M$              & Duration of maintenance. \\
        $E$              & Maximum elapsed time before assigning a maintenance. \\
        $E_{min}$        & Minimum elapsed time before assigning a maintenance. \\
        $R_{j}$          & Need of resources for task $j$. \\
        $H$              & remaining usage time after a maintenance. \\
        $D_t$            & number of resources to be assigned in total in period $t$. \\
        $A^min_j$        & minimum number of consecutive periods to assign any resource to task $j$. \\
        $Ret^{Init}_{i}$ & remaining elapsed time for resource $i$ at the start of the planning horizon. \\
    \end{tabular}

        \begin{tabular}{ll}
            $at^{init}_{it}$    &  =1 if resource $i \in I$ starts a mission assignment in period $t \in \mathcal{T}$, 0 otherwise. \\
        \end{tabular}

    The model would need to have constraints of minimum assignment. For this, we would have to implement a new variable for missions that would be similar to the one in maintenances (a 'start' of mission assignment) and then tie this new variable to the regular assignments.

        \begin{align}
            &  at^{init}_{it} \geq at_{it}
                & i \in \mathcal{I} \label{eq:min_maintenace2}\\
        \end{align}
